Working title: Highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2019-09-11 write_up/mixfishsim_ecolmod.tex
Uncommented the figures
Revisions and clean up of the tex file

\replaced[id = JJ]{Most fisheries}{Fishing} exploit\deleted[id=JJ]{s} a variety 
of spatially and temporally heterogeneous fish populations using
species-unselective gear. This can result in unintended, unwanted catch of low
quota or protected species and reducing these catches is crucial for
biological and economic sustainability of `mixed fisheries'. \\

\replaced[id = PD]{If fisheries are to avoid unwanted catch,}{To implement
	effective spatial measures to reduce discards} a good understanding of
spatiotemporal fishery dynamics is required. However, traditional scientific
advice is limited by a lack of highly resolved knowledge of population
distributions and movement, and how fishers interact with different fish
populations. Because data on fish location at high temporal and spatial
resolutions is expensive and difficult to collect proxies are usually inferred
from scientific surveys or commercial catches with limited spatial and temporal
resolution. \\

To understand how data resolution impacts inference on mixed fisheries
interactions, we developed a highly resolved spatiotemporal simulation model
incorporating: i) delay-difference population dynamics, ii) population movement
using Gaussian Random Fields to simulate patchy, heterogeneously distributed
populations, and iii) fishery dynamics for multiple fleet characteristics based
on species targeting under an explore-exploit strategy. This is implemented via
a mix of correlated random walk movement (for exploration) and learned
behaviour (for exploitation) phases of the fisheries.  \\

We simulated 50 years of fishing and used the results to draw inference on the
underlying community structures. We compared inference from commercial catch to
a simulated fixed-site sampling design and the true underlying community. We i)
establish the potential limitations of fishery-dependent data in providing a
robust picture of spatiotemporal distributions; and then ii) simulated an area
closure based on areas defined from the different data sources at a range of
temporal and spatial resolutions. We assessed the effectiveness of the closures
in reducing catches of a fish population. \\

Our framework allows users to explore the assumptions in modelling
observational data and evaluate the underlying dynamics of such approaches at a
fine spatial and temporal scale. We conclude from our simulations that
commercial data, while containing bias, provide a useful tool for managing
catches in mixed fisheries if applied at the correct spatiotemporal scale. \\

\end{abstract}

@@

Fishers exploit a variety of fish populations that are heterogeneously
distributed in space and time, with varying knowledge of species
distributions\deleted[id=JJ]{ and using species non-selective fishing gear}.
\added[id=PD]{In doing so} fisheries \deleted[id=PD]{that}catch an assemblage
of species \added[id=PD]{and}\deleted{, known as mixed fisheries}
\replaced[id=JJ]{may}{, when managed by single-species quotas can end up}
discard\deleted[id=JJ]{ing} over-quota catch \added[id=JJ]{when managed by
	single species quotas and fishers exhaust one or more quota. This may}
lead to overexploitation of fish populations \added[id=JJ]{\citep{Ulrich2011a,
		Batsleer2015}}.  \added[id=JJ]{Discarding of fish in
	excess of quota limits the ability to maintain fishing mortality within
	sustainable limits \citep{Alverson1994, Crowder1998, Rijnsdorp2007}}
\deleted[id=PD]{; reducing discarding is crucial}and the ability to manage for
the biological and economic sustainability of fisheries\deleted[id=JJ]{ and
	implementation of an ecosystem approach to fisheries}. As such, there
is increasing interest in technical solutions such as gear and spatial closures
as measures to \replaced[id=PD]{reduce unwanted catch}{avoiding
	discard\added[id=JJ]{ing of fish}} \citep{Kennelly2002, Catchpole2008,
	Bellido2011, Cosgrove2019}.  \\

\replaced[id=PD]{Adaptive spatial management strategies have}{Use of spatial
	management as a tools} been proposed as a way of reducing over-quota
discards \added[id=PD]{\citep{Holmes2011, Little2014, Dunn2014a}}.
\deleted[id=PD]{its} However, implementation is restricted by lack of knowledge
of fish and fishery spatiotemporal dynamics and understanding of the scale at
which processes become important for management. Understanding the correct
scale for spatial measures is crucial for implementing effective solutions
which  minimising economic impact \citep{Dunn2016}. For example, the problem
can be to identify a scale that promotes species avoidance for vulnerable or
low quota species while allowing continuance of sustainable fisheries for
available quota species.\\

\replaced[id=PD]{Identifying}{Ensuring measures are implemented at} appropriate
spatial scales for fisheries closures has been a challenge in the past but is
crucial to their success \citep{Costello2010, Dunn2016}. Inference is hampered
where spatial information is coarse due to low resolution reporting of
fisheries catch which is aggregated across larger gridded areas
\citep{Branch2005}. Further, if data does not allow identification of spatial
features it may lead to poorly sited closures which are ineffectual or have
unintended consequences. For example, increased benthic impact on previously
unexploited areas from the cod closure in the North Sea were observed without
the intended effect of reducing cod exploitation
\citep{Rijnsdorp2001,Dinmore2003}). \replaced[id=PD]{M}{Since then m}ore
refined spatiotemporal information has \added[id=PD]{since} become available
through the combination of logbook and Vessel Monitoring System (VMS) data
\citep{Lee2010, Bastardie2010, Gerritsen2012, Mateo2016} and more real-time
spatial management has been possible \citep[e.g.][]{Holmes2011}. However,
fishers establish favoured fishing grounds through an explore-exploit strategy
\citep{Bailey2018} where they search for areas with high catches and then use
experience to return to areas where they've experienced high catch in the past.
This leads to an inherently biased sampling where target species are
over-represented in catch as fishers exploit areas of high abundance.
\deleted[id=PD]{ Further, fishers generally only recorded landings (not catch)
	on a daily basis\todo{\added[id=JJ]{Its not clear what the problem is:
			landings collected on daily basis or landings recorded
			rather than catch}}.  This leads to questions about the
	validity of inference that can be drawn from landings data assigned to
	VMS activity pings.}\todo{\added[id=JJ]{This comes as a surprise: I
		thought this was going to be about
		discards}\added[id=PD]{Agree, have removed this to avoid
		confusion}}\\

\replaced[id=PD]{In order to understand the effect of spatiotemporal
	aggregation of data and fishery targeting we ask two fundamental
	questions regarding inference derived from observational data:}{ and
	use the results}\deleted[id=PD]{ from the fishing model:}

@@

To answer these questions we i) develop a simulation model where population
dynamics are highly-resolved in space and time by use of a Gaussian spatial
process to define suitable habitat for different populations\replaced[id=PD]{.
	Precise locations of the fish being}{ and are} known
\added[id=PD]{directly} rather than inferred from sampling or commercial
catch\added[id=PD]{, we can use the population model to validate how inference
	from fisheries-dependent and fisheries independent sampling relates to
	the real community structure in a way we could not with real data}. We
ii) compare, at different spatial and temporal aggregations, the 'real
population' distributions to samples from fisheries-dependent and fisheries
independent catches to test if these are a true reflection of the relative
density of the populations. We then iii) simulate a fishery closure to protect
a species based on different spatial and
temporal data aggregations. \\

@@

\replaced[id=PD]{An}{We developed and implemented a fully spatially explicit
	simulation model with a} \replaced{event-based simulation model of a
	hypothetical fishery was developed as a software package
	(`MixFishSim'). The modular approach enabled efficient computation by allowing
	for}{approach, where} \added[id=PD]{sub-}modules \deleted[id=PD]{are}
implemented on time-scales appropriate to capture the characteristic of the
\replaced[id=PD]{different processes}{process modelled} (Figure 1).
\added[id=PD]{The following sub-modules were included to capture the full
	system: 1) Population dynamics, 2) Recruitment dynamics, 3) Population
	movement, 4) fishery dynamics.}\\

\deleted[id=PD]{The fishing model operated on a tow-by-tow basis, while}
\replaced[id=PD]{P}{p}opulation dynamics operate on a daily
time-step\replaced[id=PD]{, while p}{. P}opulation movement occurs on a weekly
time-step, with the fishing module operating on a tow-by-tow basis (i.e.
multiple events a day).  \deleted[id=PD]{Here we describe each of the model
components; 1) Population dynamics, 2) Recruitment dynamics, 3) Population
movement dynamics, 4) fishery dynamics.}

@@

The basic population level processes were simulated using a modified two-stage
Deriso-Schnute delay difference model which models the fish populations in
terms of aggregate biomass of recruits and mature components rather than
keeping track of individuals\citep{Deriso1980, Schnute1985, Dichmont2003}.
\added[id=PD]{A daily time-step was chosen to discretise continuous population
	processes on a biologically relevant and computationally tractable
	timescale.} \deleted[id=PD]{Here,}Population biomass growth was
modelled as a function of previous recruited biomass, intrinsic population
growth and recruitment functionally linked to the adult population size.
Biomass for each cell $c$ was incremented each day $d$ as follows (the full
parameter list is detailed in Table \ref{tab: pop_variable}): 
\begin{equation}
	\begin{split}
	B_{c,d+1} = &\\
	& (1 + \rho) B_{c,d} \cdot \mathrm{e}^{-Z_{c,d}} - \rho \cdot
	\mathrm{e}^{-Z_{c,d}} \hspace{2.9cm}
	\times \\  
	& (B_{c,d-1} \cdot \mathrm{e}^{-Z_{c,d-1}} + Wt_{R-1} \cdot \alpha_{d-1} \cdot
	R_{\tilde{y}(c,y,d-1)})
	\hspace{0.4cm} + \\
	& Wt_{R} \cdot \alpha_{d} \cdot R_{\tilde{y}(c,y,d)} 
	\end{split}
\end{equation}
where $\rho$ is Brody's coefficient, shown to be equal to $\mathrm{e}^{-K}$ when $K$ is
the growth rate from a von Bertalanffy logistic growth model
\citep{Schnute1985}. $Wt_{R-1}$ is the average weight of fish prior to
recruitment, while $Wt_{R}$ is the average recruited weight. $\alpha_{d}$
represents the proportion of fish recruited during that day for the year, while
$R_{c,\tilde{y}}$ is the annual recruits in cell $c$ for year $y$. \\

Mortality $Z_{c,d}$ can be decomposed to natural mortality, $M_{c,d}$, and
fishing mortality, $F_{c,d}$, where both $M_{c,d}$ and $F_{c,d}$ are
instantaneous rates with $M_{c,d}$ fixed and $F_{c,d}$ calculated by solving
the Baranov catch equation \citep{Hilborn1992b} for $F_{c,d}$:
\begin{equation}
C_{c,d} = \frac{F_{c,d}}{F_{c,d} + M_{c,d}} * (1 - \mathrm{e}^{-(F_{c,d} + M_{c,d})})
\cdot B_{c,d}
\end{equation}
where $C_{c,d}$ is the summed catch from the fishing model across all fleets
and vessels in cell $c$ for the population during the day $d$, and $B_{c,d}$
the daily biomass for the population in the cell. Here, catch and fishing
mortality are the sum of those across all fleets and vessels, where $F_{fl, v,
	c, d, p} = E_{fl, v, c, d} \cdot Q_{fl, p} \cdot D_{c, d, p}$ with
$fl$, $v$ and $p$ the fleet, vessel and population respectively and $E$ and $Q$
fishing effort and catchability of the gear, and $D$ is the density of the
population at the location fished. \todo{\added[id=CM]{[link $F$ to effort and
		catchability - as I think we have F as an emergent property of
		the fleets rather than something we solve for (I could be wrong
		though!) - catch for a vessel is a product of catchability and
		biomass, i.e. C = qB, but this catch is summed to solve for F.
		So its both really]}}\\

@@

Recruitment is modelled through a function relating to adult biomass. In
\emph{MixFishSim}, it can be either as a stochastic Beverton-Holt stock-recruit
form \citep{Beverton1957}: 
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & \frac{(\alpha \cdot S_{c,d})}{(\beta + S_{c,d})} \\
	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\sigma^2)]
	\end{split}
\end{equation}
Where $\alpha$ is the maximum recruitment rate, $\beta$ the spawning stock
biomass (SSB) required to produce half the maximum stock size, $S$ current
stock size and $\sigma^2$ the variability in the recruitment due to stochastic
processes, or a stochastic Ricker form \citep{Ricker1954}:
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & B_{c,d} \cdot \mathrm{e}^{(\alpha - \beta \cdot B_{c,d})} \\	
   	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\log(\sigma^2))]
	\end{split}
\end{equation}
where $\alpha$ is the maximum productivity per spawner and $\beta$ the density
dependent reduction in productivity as the SSB increases.

@@

Population movement is a combination of random (diffusive) movement, governed
by a stochastic process where movement between adjacent cells is described by a
set of probabilities, and directed (advective) movement where at certain times
of year the population moves towards spawning grounds by increasing the
probabilities of moving into the spawning grounds from adjacent cells. We
incorporate characterisation of a number of different fishing
\added[id=PD]{fleet dynamics} exploiting four fish populations with different
spatial and population demographics. The following describes the implementation
of each of the sub-modules.

To simulate \deleted[id=JJ]{how} fish population\deleted[id=JJ]{s might be}
distribut\replaced[id=JJ]{ion}{ed} in space and time\deleted[id=JJ]{, we
	employed} a Gaussian spatial process \added[id=JJ]{was employed} to
model habitat suitability for each of the populations on a 2d grid. \\

\deleted[id=PD]{the habitat}We first define\added[id=PD]{d} a Gaussian random
field process\todo{\added[id=JJ]{Not clear how habitat/GRF affect local
		abundances, only have $B_{y,d}$}\added[id=PD]{Have included
		cell reference, $c$ to make spatial link explicit}}, $\{S(c) :
c \in \mathbb{R}^2\}$, \deleted[id=PD]{that is a stochastic process} where
\added[id=PD]{for} any set of cells $c_{1}, \dots, c_{n}$ \deleted[id=PD]{where
	for each $c_{i} \in \mathbb{R}^2$}, the joint distribution of $S =
\{S(c1),\dots S(c_{n})\}$ is multivariate Gaussian with a \textit{Matérn}
\deleted[id=PD]{family of}covariance structure\deleted[id=PD]{s},
\replaced[id=PD]{where}{one where} the correlation strength weakens with
distance\deleted[id=PD]{(i.e. the correlation between $S(x)$ and $S(x')$
	decreases as the distance $u = ||x - x'||$ increases)}.
\added[id=PD]{This enables us to model the spatial autocorrelation observed in
	animal populations where density is more similar in nearby locations
	\citep{Tobler1970, F.Dormann2007}} and we change the parameters to
implement different spatial structures for the different populations using the
\textit{RandomFields} R package \citep{Schlater2015}. We define a stationary
habitat field and combine with a temporally dynamic thermal tolerance field to
imitate two key drivers of population dynamics. Each population
\replaced[id=PD]{was}{is} initialised at a single location, and subsequently
move\replaced[id=PD]{d}{s} according to a probabilistic
distrib\added[id=PD]{u}tion based on habitat suitability (represented by the
normalised values from the GRFs), temperature and distance
from current cell\replaced[id=PD]{:}{.} 
\begin{equation}
	Pr(J | I) = \frac{\mathrm{e}^{-\lambda * d_{IJ}} \cdot
		(Hab_{J,p}^2 \cdot
		Tol_{J,p,wk})}{\sum\limits_{c=1}^{C}\mathrm{e}^{-\lambda * d} \cdot
		(Hab_{c,p}^2 \cdot Tol_{c,p,wk})}
\end{equation}
Where $d_{IJ}$ is the euclidean distance between cell $I$ and cell $J$,
$\lambda$ is a given rate of decay, $Hab_{J,p}^2$ is the squared index of
habitat suitability for cell $J$ and population $p$, with $Tol_{J,p,wk}$ the
temperature tolerance for cell $J$ by population $p$ in week $wk$ (see
below).\\

During pre-defined weeks of the year the habitat suitability is modified with
\added[id=PD]{user-defined} spawning habitat locations\deleted[id=PD]{s},
\replaced[id=PD]{resulting in}{meaning} each population
\replaced[id=PD]{having}{has} concentrated areas where spawning takes place. In
the simulations the populations move\deleted[id=PD]{s} towards
\replaced[id=PD]{these cells}{this} in the weeks prior to spawning, resulting
in directional movement towards the spawning grounds.\todo{\added[id=JM]{What
		does it mean concisely? Areas are assigned?} \added[id=PD]{Yes,
		the areas are pre-defined - I have amended to reflect and tried
		to clarify}} \\

\replaced[id=JJ]{An}{, with an} advection-diffusion process\deleted[id=JJ]{to}
control\added[id=JJ]{s}\deleted[id=JJ]{how the} population\deleted[id=JJ]{s}
move\replaced[id=JJ]{ment}{d}, with a time-varying temperature covariate used
to change the interaction between time and suitable habitat on a weekly
time-step\todo{\added[id=JJ]{What have a temperature covariate? Could just use
		time}\added[id=PD]{Was intended as some biological meaning -
		species thermal tolerances load onto the temperature effect -
		so could be different per species}}.  \deleted[id=JJ]{This was
	intended to balance realism in population movement, capturing the main
	directed and random processes, and practicality of modelling the
	population rather than individual fish.} Each population $p$
\replaced[id=PD]{wa}{i}s assigned a thermal tolerance with mean,
$\mu\added[id=PD]{_{p}}$ and variance, $\sigma^2\added[id=PD]{_{p}}$ so that
each cell and population temperature suitability is defined that:
\begin{equation}
	Tol_{c, p\added[id=PD]{, wk}} = \frac{1}{ \sqrt (2\pi \cdot \sigma^2_{p})} \cdot \exp(-
		\frac{(T_{c\added[id=PD]{, wk}} - \mu_{p})^2}{2 \cdot \sigma^2_{p}} )	
\end{equation}
Where $Tol_{c, p, \added[id=PD]{wk}}$ is the tolerance of population $p$ for
cell $c$ \added[id=PD]{in week $wk$}, $T_{c, \added[id=PD]{wk}}$ is the
temperature in the cell \added[id=PD]{given the week} and
$\mu\added[id=PD]{_{p}}$ and $\sigma^2\added[id=PD]{_{p}}$ the mean and
standard deviation of the population temperature tolerance. \\

\added[id=PD]{The final combined process results in a population structure and
	movement pattern unique to each species, with population movement
	occurring on a weekly basis. The decision to model population movement
	on a weekly timescale was to reflect that fish tend to aggregate in
	species specific locations that have been observed to last around one
	to two weeks \citep{Poos2007}. Therefore this process approximated the
	demographic shifts in fish populations throughout a year with seasonal
	spawning patterns (Figure \ref{S-fig:5}).}

@@

The fleet dynamics can be broadly categorised into three components; fleet
targeting - that determined the fleet catch efficiency and preference towards a
particular species; trip-level decisions, that determined the initial location
to be fished at the beginning of a trip; and within-trip decisions, that
determined movement from one fishing spot to another within a trip. Together,
these elements implemented an explore-exploit type strategy for individual
vessels to maximise their catch from an unknown resource distribution
\citep{Bailey2018}. The decision to use an individual based model for fishing
vessels was taken because fishers are heterogeneous in their location choice
behaviour due to different objectives, risk preference and targeting preference
\citep{VanPutten2012a}. Therefore in the simulations fleet dynamics are the
productive of individual experiences rather than pre-defined group dynamics.

@@

Each fleet of \textit{n} vessels \replaced[id=PD]{was}{is} characterised by
both a general efficiency, $Q_{fl}$, and a population specific efficiency,
${Q_{fl, p}}$.  Thus, the product of these parameters [$Q_{fl} \cdot Q_{fl,
	p}$] affect\replaced[id=PD]{s}{ed} the overall catch rates for the
fleet and the preferential targeting of one population over another. This, in
combination with the parameter choice for the step-function
\added[id=PD]{defined below} (as well as some randomness from the exploratory
fishing process) determine\replaced[id=PD]{d}{s} the preference of fishing
locations for the fleet.

@@

Several studies \citep[e.g.][]{Hutton2004, Tidd2012, Girardin2015} have
confirmed past activity and past catch rates are strong predictors of fishing
location choice. For this reason, the fleet dynamics sub-model
include\replaced[id=PD]{d}{s} a learning component, where a vessel's initial
fishing location in a trip \replaced[id=PD]{wa}{i}s based on selecting from
previously successful fishing locations. This \replaced[id=PD]{wa}{i}s achieved
by \added[id=PD]{calculating an expected revenue based on the catches from
	locations fished} in the preceding trip as well as the same month
periods in previous years and the travel costs from the port to the fishing
grounds, and choosing randomly from the top 75 \% of fishing events
\replaced[id=PD]{as defined by the expected profit, that has a seasonal
	component}{in value}.

@@

>>>>

Most fisheries exploit a variety of spatially and temporally heterogeneous fish
populations using species-unselective gear. This can result in unintended,
unwanted catch of low quota or protected species and reducing these catches is
crucial for biological and economic sustainability of `mixed fisheries'. \\
%%
If fisheries are to avoid unwanted catch, an in-depth understanding of
spatiotemporal fishery dynamics is required. However, traditional scientific
advice is limited by a lack of highly resolved knowledge of population
distributions and movement, and how fishers interact with different fish
populations. Because data on fish location at high temporal and spatial
resolutions is expensive and difficult to collect proxies are usually inferred
from scientific surveys or commercial catches with limited spatial and temporal
resolution. \\ 
%%
To understand how data resolution impacts inference on mixed fisheries
interactions, we developed a highly resolved spatiotemporal
event-based simulation model incorporating: i) delay-difference
population dynamics, ii) population movement using Gaussian Random Fields to
simulate patchy, heterogeneously distributed populations, and iii) fishery
dynamics for multiple fleet characteristics based on species targeting under an
explore-exploit strategy. This is implemented via a mix of correlated random
walk movement (for exploration) and learned behaviour (for exploitation) phases
of the fisheries.  \\ 
%%
Fifty years of sub-daily fishing was simulated and used to draw inference on
the underlying community structures. We compared inferences based on:
commercial catch, a simulated fixed-site sampling survey design and the true
underlying community. We i) establish the potential limitations of
fishery-dependent data in providing a robust picture of spatiotemporal
distributions; and then ii) simulated an area closure based on areas defined
from the different data sources at a range of temporal and spatial resolutions.
\\
%%
Our framework allows users to explore the assumptions in modelling
observational data and evaluate the underlying dynamics of such approaches at a
fine spatial and temporal scale. We conclude from our example framework
application that commercial data, while containing bias, provide a useful tool
for managing catches in mixed fisheries if applied at the correct
spatiotemporal scale. [NEED A LITTLE MORE DETAIL OF THE RESULTS HERE - 2 MORE
SENTENCES ON KEY TAKE-HOME RESULTS]\\
%%
\end{abstract}

@@

Fishers exploit a variety of fish populations that are heterogeneously
distributed in space and time, with varying knowledge of species distributions.
In doing so fisheries catch an assemblage of species and may discard over-quota
catch when managed by single species quotas and fishers exhaust one or more
quota. This may lead to overexploitation of fish populations
\citep{Ulrich2011a, Batsleer2015}. Discarding of fish in excess of quota limits
the ability to maintain fishing mortality within sustainable limits
\citep{Alverson1994, Crowder1998, Rijnsdorp2007} and the ability to manage for
the biological and economic sustainability of fisheries. As such, there is
increasing interest in technical solutions such as gear and spatial closures as
measures to reduce unwanted catch \citep{Kennelly2002, Catchpole2008,
	Bellido2011, Cosgrove2019}. \\

Adaptive spatial management strategies have been proposed as a way of reducing
over-quota discards \citep{Holmes2011, Little2014, Dunn2014a}. However,
implementation is restricted by lack of knowledge of fish and fishery
spatiotemporal dynamics and understanding of the scale at which processes
become important for management. Understanding the correct scale for spatial
measures is crucial for implementing effective solutions which  minimising
economic impact \citep{Dunn2016}. For example, the problem can be to identify a
scale that promotes species avoidance for vulnerable or low quota species while
allowing continuance of sustainable fisheries for available quota species.\\

Identifying appropriate spatial scales for fisheries closures has been a
challenge in the past but is crucial to their success \citep{Costello2010,
	Dunn2016}. Inference is hampered where spatial information is coarse
due to low resolution reporting of fisheries catch which is aggregated across
larger gridded areas \citep{Branch2005}. Further, if data does not allow
identification of spatial features it may lead to poorly sited closures which
are ineffectual or have unintended consequences. For example, increased benthic
impact on previously unexploited areas from the cod closure in the North Sea
were observed without the intended effect of reducing cod exploitation
\citep{Rijnsdorp2001,Dinmore2003}). More refined spatiotemporal information has
since become available through the combination of logbook and Vessel Monitoring
System (VMS) data \citep{Lee2010, Bastardie2010, Gerritsen2012, Mateo2016} and
more real-time spatial management has been possible \citep[e.g.][]{Holmes2011}.
However, fishers establish favoured fishing grounds through an explore-exploit
strategy \citep{Bailey2018} where they search for areas with high catches and
then use experience to return to areas where they've experienced high catch in
the past.  This leads to an inherently biased sampling where target species are
over-represented in the catch as fishers exploit areas of high abundance. \\

To understand the effect of spatiotemporal aggregation of data and fishery
targeting we ask two fundamental questions regarding inference derived from
observational data:

@@

To answer these questions we i) develop a simulation model where population
dynamics are highly-resolved in space and time by use of a Gaussian spatial
process to define suitable habitat for different populations.  Precise
locations of the fish being known directly rather than inferred from sampling
or commercial catch, we can use the population model to validate how inference
from fisheries-dependent and fisheries independent sampling relates to the real
community structure in a way we could not with real data. We ii) compare, at
different spatial and temporal aggregations, the real (simulated) population
distributions to samples from fisheries-dependent and fisheries independent
catches to test if these are a true reflection of the relative density of the
populations. We then iii) simulate a fishery closure to protect a species based
on different spatial and temporal data aggregations. \\

@@

An event-based simulation model of a hypothetical fishery was developed as a
software package (\textit{MixFishSim}). The modular approach enabled efficient
computation by allowing for sub-modules implemented on time-scales appropriate
to capture the characteristic of the different processes (Figure 1). The
following sub-modules were included to capture the full system: 1) Population
dynamics, 2) Recruitment dynamics, 3) Population movement, 4) fishery
dynamics.\\

Population dynamics operate on a daily time-step, while population movement
occurs on a weekly time-step, with the fishing module operating on a tow-by-tow
basis (i.e.  multiple events a day).

@@

The basic population level processes were simulated using a modified two-stage
Deriso-Schnute delay difference model which models the fish populations in
terms of aggregate biomass of recruits and mature components rather than
keeping track of individuals\citep{Deriso1980, Schnute1985, Dichmont2003}. A
daily time-step was chosen to discretise continuous population processes on a
biologically relevant and computationally tractable timescale. Population
biomass growth was modelled as a function of previous recruited biomass,
intrinsic population growth and recruitment functionally linked to the adult
population size.  Biomass for each cell $c$ was incremented each day $d$ as
follows (the full parameter list is detailed in Table \ref{tab: pop_variable}): 
\begin{equation}
	\begin{split}
	B_{c,d+1} = &\\
	& (1 + \rho) B_{c,d} \cdot \mathrm{e}^{-Z_{c,d}} - \rho \cdot
	\mathrm{e}^{-Z_{c,d}} \hspace{2.9cm}
	\times \\  
	& (B_{c,d-1} \cdot \mathrm{e}^{-Z_{c,d-1}} + Wt_{R-1} \cdot \alpha_{d-1} \cdot
	R_{\tilde{y}(c,y,d-1)})
	\hspace{0.4cm} + \\
	& Wt_{R} \cdot \alpha_{d} \cdot R_{\tilde{y}(c,y,d)} 
	\end{split}
\end{equation}
where $\rho$ is Brody's coefficient, shown to be equal to $\mathrm{e}^{-K}$
when $K$ is the growth rate from a von Bertalanffy logistic growth model
\citep{Schnute1985}. $Wt_{R-1}$ is the average weight of fish prior to
recruitment, while $Wt_{R}$ is the average recruited weight. $\alpha_{d}$
represents the proportion of fish recruited during that day for the year, while
$R_{c,\tilde{y}}$ is the annual recruits in cell $c$ for year $y$. \\

Mortality $Z_{c,d}$ can be decomposed to natural mortality, $M_{c,d}$, and
fishing mortality, $F_{c,d}$, where both $M_{c,d}$ and $F_{c,d}$ are
instantaneous rates with $M_{c,d}$ fixed and $F_{c,d}$ calculated by solving
the Baranov catch equation \citep{Hilborn1992b} for $F_{c,d}$:
\begin{equation}
C_{c,d} = \frac{F_{c,d}}{F_{c,d} + M_{c,d}} \cdot \left(1 -
	\mathrm{e}^{-(F_{c,d} + M_{c,d})}\right) \cdot B_{c,d}
\end{equation}
where $C_{c,d}$ is the summed catch from the fishing model across all fleets
and vessels in cell $c$ for the population during the day $d$, and $B_{c,d}$
the daily biomass for the population in the cell. Here, catch and fishing
mortality are the sum of those across all fleets and vessels, where $F_{fl, v,
	c, d, p} = E_{fl, v, c, d} \cdot Q_{fl, p} \cdot D_{c, d, p}$ with
$fl$, $v$ and $p$ the fleet, vessel and population respectively and $E$ and $Q$
fishing effort and catchability of the gear, and $D$ is the density of the
population at the location fished. \\

@@

Recruitment is modelled through a function relating to adult biomass. In
\textit{MixFishSim}, it can be either as a stochastic Beverton-Holt stock-recruit
form \citep{Beverton1957}: 
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & \frac{(\alpha \cdot S_{c,d})}{(\beta + S_{c,d})} \\
	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\sigma^2)]
	\end{split}
\end{equation}
Where $\alpha$ is the maximum recruitment rate, $\beta$ the spawning stock
biomass (SSB) required to produce half the maximum stock size, $S$ current
stock size and $\sigma^2$ the variability in the recruitment due to stochastic
processes, or a stochastic Ricker form \citep{Ricker1954}:
\begin{equation}
	\begin{split}
	\bar{R}_{c,d} = & B_{c,d} \cdot \mathrm{e}^{(\alpha - \beta \cdot B_{c,d})} \\	
   	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\log(\sigma^2))]
	\end{split}
\end{equation}
where $\alpha$ is the maximum productivity per spawner and $\beta$ the density
dependent reduction in productivity as the SSB increases.

@@

Population movement is a combination of random (diffusive) movement, governed
by a stochastic process where movement between adjacent cells is described by a
set of probabilities, and directed (advective) movement where at certain times
of year the population moves towards spawning grounds by increasing the
probabilities of moving into the spawning grounds from adjacent cells. We
characterise a set of different fishing fleet dynamics exploiting four fish
populations with different spatial and population demographics. \\

To simulate fish population distribution in space and time a Gaussian spatial
process was employed to model habitat suitability for each of the populations
on a 2d grid.  We first defined a Gaussian random field process, $\{S(c) : c \in
\mathbb{R}^2\}$, where for any set of cells $c_{1}, \dots, c_{n}$, the joint
distribution of $S = \{S(c1),\dots S(c_{n})\}$ is multivariate Gaussian with a
\textit{Matérn} covariance structure, where the correlation strength weakens
with distance. This enables us to model the spatial autocorrelation observed in
animal populations where density is more similar in nearby locations
\citep{Tobler1970, F.Dormann2007} and we change the parameters to implement
different spatial structures for the different populations using the
\textit{RandomFields} R package \citep{Schlater2015}. We define a stationary
habitat field and combine with a temporally dynamic thermal tolerance field to
imitate two key drivers of population dynamics. Each population was initialised
at a single location, and subsequently moved according to a probabilistic
distribution based on habitat suitability (represented by the normalised values
from the GRFs), temperature and distance from current cell: 
\begin{equation}
	Pr(C_{wk+1} = J | C_{wk} =I) = \frac{\mathrm{e}^{-\lambda * d_{IJ}} \cdot
		(Hab_{J,p}^2 \cdot
		Tol_{J,p,wk})}{\sum\limits_{c=1}^{C}\mathrm{e}^{-\lambda * d} \cdot
		(Hab_{c,p}^2 \cdot Tol_{c,p,wk})}
\end{equation}
Where $d_{IJ}$ is the euclidean distance between cell $I$ and cell $J$,
$\lambda$ is a given rate of decay, $Hab_{J,p}^2$ is the squared index of
habitat suitability for cell $J$ and population $p$, with $Tol_{J,p,wk}$ the
temperature tolerance for cell $J$ by population $p$ in week $wk$ (see
below). \\

During pre-defined weeks of the year the habitat suitability is modified with
user-defined spawning habitat locations, resulting in each population having
concentrated areas where spawning takes place. In the simulations the
populations move towards these cells in the weeks prior to spawning, resulting
in directional movement towards the spawning grounds. \\

An advection-diffusion process controls population movement, with a
time-varying temperature covariate used to change the interaction between time
and suitable habitat on a weekly time-step. Each population $p$ was assigned a
thermal tolerance with mean, $\mu_{p}$ and variance, $\sigma^2_{p}$ so that
each cell and population temperature suitability is defined that:
\begin{equation}
	Tol_{c, p, wk} = \frac{1}{ \sqrt (2\pi \cdot \sigma^2_{p})} \cdot \exp(-
		\frac{(T_{c\added[id=PD]{, wk}} - \mu_{p})^2}{2 \cdot \sigma^2_{p}} )	
\end{equation}
Where $Tol_{c, p, wk}$ is the tolerance of population $p$ for cell $c$ in week
$wk$, $T_{c, wk}$ is the temperature in the cell given the week and $\mu_{p}$
and $\sigma^2_{p}$ the mean and standard deviation of the population
temperature tolerance. \\

The final combined process results in a population structure and movement
pattern unique to each species, with population movement occurring on a weekly
basis. The decision to model population movement on a weekly timescale was to
reflect that fish tend to aggregate in species specific locations that have
been observed to last around one to two weeks \citep{Poos2007}. Therefore this
process approximated the demographic shifts in fish populations throughout a
year with seasonal spawning patterns (Figure \ref{S-fig:5}).

@@

Fleet dynamics can be broadly categorised into three components: fleet
targeting - that determined the fleet catch efficiency and preference towards a
particular species; trip-level decisions, that determines the initial location
to be fished at the beginning of a trip; and within-trip decisions, that
determines movement from one fishing spot to another within a trip. Together,
these elements implement an explore-exploit type strategy for individual
vessels to maximise their catch from an unknown resource distribution
\citep{Bailey2018}. The decision to use an individual based model for fishing
vessels was taken because fishers are heterogeneous in their location choice
behaviour due to different objectives, risk preference and targeting preference
\citep{VanPutten2012a}. Therefore in the simulations fleet dynamics reflect
individual experiences rather than pre-defined group dynamics.

@@

Each fleet of \textit{n} vessels was characterised by both a general
efficiency, $Q_{fl}$, and a population specific efficiency, ${Q_{fl, p}}$.
Thus, the product of these parameters [$Q_{fl} \cdot Q_{fl, p}$] affects the
overall catch rates for the fleet and the preferential targeting of one
population over another. This, in combination with the parameter choice for the
step-function defined below (as well as some randomness from the exploratory
fishing process) determined the preference of fishing locations for the fleet.

@@

Several studies \citep[e.g.][]{Hutton2004, Tidd2012, Girardin2015} have
confirmed past activity and past catch rates are strong predictors of fishing
location choice. For this reason, the fleet dynamics sub-model included a
learning component, where a vessel's initial fishing location in a trip was
based on selecting from previously successful fishing locations. This was
achieved by calculating an expected revenue based on the catches from locations
fished in the preceding trip as well as the same month periods in previous
years and the travel costs from the port to the fishing grounds, and choosing
randomly from the top 75 \% of fishing events as defined by the expected
profit, that has a seasonal component.

@@

\subsection{Emergent simulation dynamics}
